---
title: "Computational Modeling - Week 4 - Assignment 2 - Part 1"
author: "Riccardo Fusaroli"
date: "2/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## In this assignment we learn how to assess rates from a binomial distribution, using the case of assessing your teachers' knowledge of CogSci

### First part

You want to assess your teachers' knowledge of cognitive science. "These guys are a bunch of drama(turgist) queens, mindless philosophers, chattering communication people and Russian spies. Do they really know CogSci?", you think.

To keep things simple (your teachers should not be faced with too complicated things):
- You created a pool of equally challenging questions on CogSci
- Each question can be answered correctly or not (we don't allow partially correct answers, to make our life simpler).
- Knowledge of CogSci can be measured on a scale from 0 (negative knowledge, all answers wrong) through 0.5 (random chance) to 1 (awesome CogSci superpowers)

This is the data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

```{r}
d <- data.frame(
  Correct=c(3,2,160,66),
  Questions=c(6,2,198,132),
  Teacher=c("RF","KT","JS","MW"))
```


Questions:

1. What's Riccardo's estimated knowledge of CogSci? What is the probability he knows more than chance (0.5) [try figuring this out. if you can't peek into chapters 3.1 and 3.2 and/or the slides]?
- First implement a grid approximation (hint check paragraph 2.4.1!) with a uniform prior, calculate the posterior and plot the results

```{r}
#Define the grid
dens=50 #how many points do I want?
p_grid <- seq( from=0 , to=1 , length.out=dens) #creates the grid
#Define the prior 
prior <- rep( 1 , dens ) # Flat 
# prior <- ifelse(p_grid < Ϭ.5 , Ϭ , ϭ) #informed
# prior <- exp(-5*abs( p_grid - 0.5) #conservative 
```


```{r}

#Riccardo

#Compute the likelihood at each value in grid 
#it's a binomial distribution since the answers can be right or wrong
#dbinom(correct answers(p), all answers(n), prob= go through the grid)
likelihood <- dbinom( d$Correct[d$Teacher == "RF"] , size= d$Questions [d$Teacher == "RF"] , prob=p_grid ) 
#Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior 
#Standardize the posterior (so it sums to 1) 
posterior <- unstd.posterior / sum(unstd.posterior)
#Draw the plot 
Data_R=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
Data_R$teacher= c("Riccardo")

library(ggplot2)
ggplot(Data_R,aes(grid,posterior))+ 
  geom_point()+geom_line()+
  theme_classic()+
  geom_line(aes(grid,prior/dens), color='red')+
  labs(x="probability of correct answer", y="posterior probability", title= "Riccardo")
  

#What is the probability he knows more than chance? 
#for how many values of the posterior is the number in the grid bigger than 0.5 (???)
#how does this work?
sum(posterior[p_grid > 0.5])
```

- Then implement a quadratic approximation (hint check paragraph 2.4.2!).
```{r}

library(rethinking)
globe.qa <- map(
  alist(
    w ~ dbinom(d$Questions [d$Teacher == "RF"],p) , # binomial likelihood
    p ~ dunif(0,1) # uniform prior
  ) ,
  data=list(w= d$Correct[d$Teacher == "RF"]) )
# display summary of quadratic approximation
precis( globe.qa )

```

- N.B. for the rest of the exercise just keep using the grid approximation (we'll move to quadratic approximations in two classes)

2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel.
2a. Produce plots of the prior, and posterior for each teacher. 
```{r}

#Josh

likelihood <- dbinom( d$Correct[d$Teacher == "JS"] , size= d$Questions [d$Teacher == "JS"] , prob=p_grid ) 
#Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior 
#Standardize the posterior (so it sums to 1) 
posterior <- unstd.posterior / sum(unstd.posterior)
#Draw the plot 
Data_J=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
Data_J$teacher= c("Josh")

ggplot(Data_J,aes(grid,posterior))+ 
  geom_point()+geom_line()+
  theme_classic()+
  geom_line(aes(grid,prior/dens), color='red')+
  labs(x="probability of correct answer", y="posterior probability", title= "Josh")


#Mikkel

likelihood <- dbinom( d$Correct[d$Teacher == "MW"] , size= d$Questions [d$Teacher == "MW"] , prob=p_grid ) 
#Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior 
#Standardize the posterior (so it sums to 1) 
posterior <- unstd.posterior / sum(unstd.posterior)
#Draw the plot 
Data_M=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
Data_M$teacher= c("Mikkel")

ggplot(Data_M,aes(grid,posterior))+ 
  geom_point()+geom_line()+
  theme_classic()+
  geom_line(aes(grid,prior/dens), color='red')+
  labs(x="probability of correct answer", y="posterior probability", title= "Mikkel")

#Kristian

likelihood <- dbinom( d$Correct[d$Teacher == "KT"] , size= d$Questions [d$Teacher == "KT"] , prob=p_grid ) 
#Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior 
#Standardize the posterior (so it sums to 1) 
posterior <- unstd.posterior / sum(unstd.posterior)
#Draw the plot 
Data_K=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
Data_K$teacher= c("Kristian")


ggplot(Data_K,aes(grid,posterior))+ 
  geom_point()+geom_line()+
  theme_classic()+
  geom_line(aes(grid,prior/dens), color='red')+
  labs(x="probability of correct answer", y="posterior probability", title= "Kristian")

data_all= rbind(Data_R, Data_J, Data_M, Data_K)

ggplot(data_all,aes(grid,posterior, colour= teacher))+ 
  geom_line()+
  theme_classic()+
  geom_line(aes(grid,prior/dens), color='red')+
  labs(x="probability of correct answer", y="posterior probability", title= "all teachers")

```


3. Change the prior. Given your teachers have all CogSci jobs, you should start with a higher appreciation of their knowledge: the prior is a normal distribution with a mean of 0.8 and a standard deviation of 0.2. Do the results change (and if so how)?
3a. Produce plots of the prior and posterior for each teacher.

```{r}

prior = dnorm(p_grid, mean= 0.8, sd= 0.2)

#just running again previous code
#Riccardo

likelihood <- dbinom( d$Correct[d$Teacher == "RF"] , size= d$Questions [d$Teacher == "RF"] , prob=p_grid ) 
unstd.posterior <- likelihood * prior 
posterior <- unstd.posterior / sum(unstd.posterior)
Data_R=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
Data_R$teacher= c("Riccardo")

ggplot(Data_R,aes(grid,posterior))+ 
  geom_point()+geom_line()+
  theme_classic()+
  geom_line(aes(grid,prior/dens), color='red')+
  labs(x="probability of correct answer", y="posterior probability", title= "Riccardo")
  

#Josh

likelihood <- dbinom( d$Correct[d$Teacher == "JS"] , size= d$Questions [d$Teacher == "JS"] , prob=p_grid ) 
unstd.posterior <- likelihood * prior 
posterior <- unstd.posterior / sum(unstd.posterior)
Data_J=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
Data_J$teacher= c("Josh")

ggplot(Data_J,aes(grid,posterior))+ 
  geom_point()+geom_line()+
  theme_classic()+
  geom_line(aes(grid,prior/dens), color='red')+
  labs(x="probability of correct answer", y="posterior probability", title= "Josh")


#Mikkel

likelihood <- dbinom( d$Correct[d$Teacher == "MW"] , size= d$Questions [d$Teacher == "MW"] , prob=p_grid ) 
unstd.posterior <- likelihood * prior 
posterior <- unstd.posterior / sum(unstd.posterior)
Data_M=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
Data_M$teacher= c("Mikkel")

ggplot(Data_M,aes(grid,posterior))+ 
  geom_point()+geom_line()+
  theme_classic()+
  geom_line(aes(grid,prior/dens), color='red')+
  labs(x="probability of correct answer", y="posterior probability", title= "Mikkel")

#Kristian

likelihood <- dbinom( d$Correct[d$Teacher == "KT"] , size= d$Questions [d$Teacher == "KT"] , prob=p_grid ) 
unstd.posterior <- likelihood * prior 
posterior <- unstd.posterior / sum(unstd.posterior)
Data_K=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
Data_K$teacher= c("Kristian")


ggplot(Data_K,aes(grid,posterior))+ 
  geom_point()+geom_line()+
  theme_classic()+
  geom_line(aes(grid,prior/dens), color='red')+
  labs(x="probability of correct answer", y="posterior probability", title= "Kristian")

data_all= rbind(Data_R, Data_J, Data_M, Data_K)

ggplot(data_all,aes(grid,posterior, colour= teacher))+ 
  geom_line()+
  theme_classic()+
  geom_line(aes(grid,prior/dens), color='red')+
  labs(x="probability of correct answer", y="posterior probability", title= "all teachers")

```


4. You go back to your teachers and collect more data (multiply the previous numbers by 100). Calculate their knowledge with both a uniform prior and a normal prior with a mean of 0.8 and a standard deviation of 0.2. Do you still see a difference between the results (between using the different priors)? Why? (reflect how much the data and priors matter)

```{r}
d2 <- data.frame(
  Correct=c(3,2,160,66)*100,
  Questions=c(6,2,198,132)*100,
  Teacher=c("RF","KT","JS","MW"))

#Define the grid
dens=50 #how many points do I want?
p_grid <- seq( from=0 , to=1 , length.out=dens) #creates the grid
#Define the prior 

prior <- rep( 1 , dens )

#just running again previous code

#Riccardo

likelihood <- dbinom( d2$Correct[d2$Teacher == "RF"] , size= d2$Questions [d2$Teacher == "RF"] , prob=p_grid ) 
unstd.posterior <- likelihood * prior 
posterior <- unstd.posterior / sum(unstd.posterior)
Data_R=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
Data_R$teacher= c("Riccardo")

ggplot(Data_R,aes(grid,posterior))+ 
  geom_point()+geom_line()+
  theme_classic()+
  geom_line(aes(grid,prior/dens), color='red')+
  labs(x="probability of correct answer", y="posterior probability", title= "Riccardo")
  

#Josh

likelihood <- dbinom( d2$Correct[d2$Teacher == "JS"] , size= d2$Questions [d2$Teacher == "JS"] , prob=p_grid ) 
unstd.posterior <- likelihood * prior 
posterior <- unstd.posterior / sum(unstd.posterior)
Data_J=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
Data_J$teacher= c("Josh")

ggplot(Data_J,aes(grid,posterior))+ 
  geom_point()+geom_line()+
  theme_classic()+
  geom_line(aes(grid,prior/dens), color='red')+
  labs(x="probability of correct answer", y="posterior probability", title= "Josh")


#Mikkel

likelihood <- dbinom( d2$Correct[d2$Teacher == "MW"] , size= d2$Questions [d2$Teacher == "MW"] , prob=p_grid ) 
unstd.posterior <- likelihood * prior 
posterior <- unstd.posterior / sum(unstd.posterior)
Data_M=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
Data_M$teacher= c("Mikkel")

ggplot(Data_M,aes(grid,posterior))+ 
  geom_point()+geom_line()+
  theme_classic()+
  geom_line(aes(grid,prior/dens), color='red')+
  labs(x="probability of correct answer", y="posterior probability", title= "Mikkel")

#Kristian

likelihood <- dbinom( d2$Correct[d2$Teacher == "KT"] , size= d2$Questions [d2$Teacher == "KT"] , prob=p_grid ) 
unstd.posterior <- likelihood * prior 
posterior <- unstd.posterior / sum(unstd.posterior)
Data_K=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
Data_K$teacher= c("Kristian")


ggplot(Data_K,aes(grid,posterior))+ 
  geom_point()+geom_line()+
  theme_classic()+
  geom_line(aes(grid,prior/dens), color='red')+
  labs(x="probability of correct answer", y="posterior probability", title= "Kristian")




```

```{r}
prior = dnorm(p_grid, mean= 0.8, sd= 0.2)

#just running again previous code

#Riccardo

likelihood <- dbinom( d2$Correct[d2$Teacher == "RF"] , size= d2$Questions [d2$Teacher == "RF"] , prob=p_grid ) 
unstd.posterior <- likelihood * prior 
posterior <- unstd.posterior / sum(unstd.posterior)
Data_R=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
Data_R$teacher= c("Riccardo")

ggplot(Data_R,aes(grid,posterior))+ 
  geom_point()+geom_line()+
  theme_classic()+
  geom_line(aes(grid,prior/dens), color='red')+
  labs(x="probability of correct answer", y="posterior probability", title= "Riccardo")
  

#Josh

likelihood <- dbinom( d2$Correct[d2$Teacher == "JS"] , size= d2$Questions [d2$Teacher == "JS"] , prob=p_grid ) 
unstd.posterior <- likelihood * prior 
posterior <- unstd.posterior / sum(unstd.posterior)
Data_J=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
Data_J$teacher= c("Josh")

ggplot(Data_J,aes(grid,posterior))+ 
  geom_point()+geom_line()+
  theme_classic()+
  geom_line(aes(grid,prior/dens), color='red')+
  labs(x="probability of correct answer", y="posterior probability", title= "Josh")


#Mikkel

likelihood <- dbinom( d2$Correct[d2$Teacher == "MW"] , size= d2$Questions [d2$Teacher == "MW"] , prob=p_grid ) 
unstd.posterior <- likelihood * prior 
posterior <- unstd.posterior / sum(unstd.posterior)
Data_M=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
Data_M$teacher= c("Mikkel")

ggplot(Data_M,aes(grid,posterior))+ 
  geom_point()+geom_line()+
  theme_classic()+
  geom_line(aes(grid,prior/dens), color='red')+
  labs(x="probability of correct answer", y="posterior probability", title= "Mikkel")

#Kristian

likelihood <- dbinom( d2$Correct[d2$Teacher == "KT"] , size= d2$Questions [d2$Teacher == "KT"] , prob=p_grid ) 
unstd.posterior <- likelihood * prior 
posterior <- unstd.posterior / sum(unstd.posterior)
Data_K=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
Data_K$teacher= c("Kristian")


ggplot(Data_K,aes(grid,posterior))+ 
  geom_point()+geom_line()+
  theme_classic()+
  geom_line(aes(grid,prior/dens), color='red')+
  labs(x="probability of correct answer", y="posterior probability", title= "Kristian")

```


5. Imagine you're a skeptic and think your teachers do not know anything about CogSci, given the content of their classes. How would you operationalize that belief?

6. Optional question: Can you estimate the difference between Riccardo's estimated knowledge and that of each of the other teachers? Would you deem it credible (that is, would you believe that it is actually different)?

7. Bonus knowledge: all the stuff we have done can be implemented in a lme4-like fashion using the brms package. Here is an example.
```{r}
# library(brms)
# d <- data.frame(
#   Correct=c(3,2,160,66),
#   Questions=c(6,2,198,132),
#   Teacher=c("RF","KT","JS","MW"))
# 
# FlatModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("uniform(0,1)", class = "Intercept"))
# plot(FlatModel)
# PositiveModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.8,0.2)", class = "Intercept"))
# plot(PositiveModel)
# SkepticalModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.5,0.01)", class = "Intercept"))
# plot(SkepticalModel)
```

If you dare, try to tweak the data and model to test two hypotheses:
- Is Kristian different from Josh?
- Is Josh different from chance?


```{r}
# #Define the grid
# dens=20 
# p_grid <- seq( from=0 , to=1 , length.out=dens) 
# #Define the prior 
# prior <- rep( 1 , dens ) # Flat 
# # prior <- ifelse(p_grid < Ϭ.5 , Ϭ , ϭ) #informed
# # prior <- exp(-5*abs( p_grid - 0.5) #conservative 
# #Compute the likelihood at each value in grid 
# likelihood <- dbinom( 6 , size=9 , prob=p_grid ) 
# #Compute the posterior (likelihood by prior)
# unstd.posterior <- likelihood * prior 
# #Standardize the posterior (so it sums to 1) 
# posterior <- unstd.posterior / sum(unstd.posterior)
# #Draw the plot 
# Data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood) 
# ggplot(Data,aes(grid,posterior))+ 
#   geom_point()+geom_line()+
#   theme_classic()+
#   geom_line(aes(grid,prior/dens), color='red')+
#   xlab("probability of water")+ ylab("posterior probability")

```



